---
title: "NBA ELO Rating"
output: html_notebook
---

# Loading and Installing Libraries
```{r}
install.packages("tidyverse")
library(tidyverse)
```

# Importing Data set NBA_ELO

```{r}
nba_ELO<-read_csv("nbaallelo.csv")
nba_ELO
```
# Choosing 2 teams from 2014-2015 NBA Finals (Warriors vs Cavaliers)
```{r}
# Filtering year 2015
nba2015<-filter(nba_ELO, year_id==2015)
nba2015

# Filtering teams to make comparisons
teams<-filter(nba2015, team_id=="gsw" & opp_id=="cle")
teams
```
# Comparing ELO ratings for both teams and their outcomes based on Game 1-4
```{r}
games<-filter(teams, date_game=="6/4/2015" | date_game=="6/7/2015" | date_game=="6/9/2015" | date_game=="6/11/2015" )
selectingdata<-select(games, fran_id, pts, elo_i, date_game,win_equiv, opp_fran, opp_pts,opp_elo_i )
selectingdata
```
# Histogram of Warriors ELO entering Game 1-4
```{r}
ggplot(data=games)+
  geom_histogram(mapping=aes(x=elo_i, fill=date_game), bins=10)+
  labs(title="2014-2015 NBA Finals Game 1-4 Results",
         subtitle="Warriors",
         x="ELO Entering Game", y="Amount")
```
# Histogram of Cavaliers ELO entering Game 1-4
```{r}
ggplot(data=games)+
  geom_histogram(mapping=aes(x=opp_elo_i, fill=date_game), bins=10)+
  labs(title="2014-2015 NBA Finals Game 1-4 Results",
         subtitle="Cavaliers",
         x="ELO Entering Game", y="Amount")
```

# Scatter plot of Warriors ELO vs Cavaliers ELO entering game 1-4
```{r}
ggplot(data=games)+
  geom_point(mapping=aes(x=opp_elo_i,y= elo_i, color=date_game))+
  labs(title="2014-2015 NBA Finals Game 1-4 Results",
         subtitle="Cavaliers vs Warriors",
         x="ELO Entering Game (cavs)", y="ELO Entering Game (Warriors)")
```

# Comparing ELO ratings during the regular season games (Cavs vs. Warriors)
```{r}
regular<-filter(teams, date_game=="1/9/2015" | date_game=="2/26/2015" )
regData<-select(regular, fran_id, pts, elo_i, date_game,win_equiv, opp_fran, opp_pts,opp_elo_i )
regData
```

# Scatter plot of Regular Season games
```{r}
ggplot(data=regular)+
  geom_point(mapping=aes(x=opp_elo_i,y= elo_i, color=date_game))+
  labs(title="2014-2015 NBA Regular Season Results",
         subtitle="Cavaliers vs Warriors",
         x="ELO Entering Game (cavs)", y="ELO Entering Game (Warriors)")
```
# ELO rating based on Game 5-6 of NBA Finals

```{r}
# filtering data
lastFinal<-filter(teams, date_game=="6/14/2015" | date_game=="6/16/2015" )
lastF<-select(lastFinal, fran_id, pts, elo_i, date_game,win_equiv, opp_fran, opp_pts,opp_elo_i )
lastF
```
# Scatter plot of Game 5-6
```{r}
ggplot(data=lastFinal)+
  geom_point(mapping=aes(x=opp_elo_i,y= elo_i, color=date_game))+
  labs(title="2014-2015 NBA Finals Game 5-6 Results",
         subtitle="Cavaliers vs Warriors",
         x="ELO Entering Game (cavs)", y="ELO Entering Game (Warriors)")
```